---
title: <center> ECS 132 Fall 2020, Project</center>
header-includes:
- \usepackage{tikz}
- \usepackage{amsmath}
- \usepackage{graphicx}
- \usepackage{fancyhdr}
- \usepackage{geometry}
- \usepackage{float}
- \usepackage{xcolor}
date: "<center> December 1, 2020 </center>"
output:
  html_document:
    
    highlight: tango
    number_sections: yes
    theme: united
    toc: no
    toc_depth: 3
    pandoc_args:
    - --number-sections
    - --number-offset=2
  pdf_document:
    toc: no
    toc_depth: '3'
    highlight: tango
    number_sections: yes
---


\pagestyle{fancy}
\fancyhf{}
\rhead{Project}
\lhead{ECS 132 Fall 2020}
\rfoot{Page \thepage}

<!-- regular html comment --> 
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```
## Your name:  Gabriel Vazquez <!-- write your name --> {-}

## Partner's name:  Neema Nejad <!-- write your name --> {-}

## Any other collaborator(s):   <!-- write your name --> {-}


# Instructions {-}

1. Please refer to "ECS_132_Project.pdf" for details.

2. You may in no circumstances upload your project to private tutoring websites such as CourseHero or Chegg. Remember all material related to this course  is a property of the University of California and posting them is a violation of the copyright laws. 
    
3. If you refer to a source (either a book or the internet), you must cite it.
    
4. You may work in groups of size of at most 2. If you collaborate with others,  you must list their names.
    
5. Write your answers in R Markdown and submit the knitted pdf on Gradescope; for due date and other details see the Homework Policy and Schedule.

\setcounter{section}{2}
# Design

## 

Alice and Bob decide to use the following modulation scheme to map the bits to the inter-packet delay. A delay of $0.25$ is used to encode a bit $0$ and delay of $0.75$ is used to encode a bit $1$. Write a short R code  that will generate the modified packet stream that contains the secret message.

## Answer {-}
```{r }
packetTimes = read.csv("Traffic_data_orig.csv")[2];
secretMsg = "this is a secret message";
binaryNumbers = numeric(length=(8*nchar(secretMsg)));

# convert character to binary numbers and store them in list.
for (i in 1:nchar(secretMsg)) {
  c = substr(secretMsg, i, i);
  binary = rev(rawToBits(charToRaw(c))); # rawToBits returns the binary number with LSB first so need to reverse it.
  for (j in 1:length(binary)) { 
    binaryNumbers[8*(i-1)+j] = if(binary[j] == 0) 0 else 1;
  }
}

modifiedPacketStream = numeric(8*nchar(secretMsg)+1);
time = 0;
modifiedPacketStream[1] = time;

for (i in 1:(length(binaryNumbers))) {
  time = time + if(binaryNumbers[i] == 0) 0.25 else 0.75;
  modifiedPacketStream[i+1] = time;
}

print(modifiedPacketStream);

```

## 

Plot the histogram of the inter-packet delays of the overt packet stream. Plot the histogram of the covert packet stream. Will Eve be suspicious?

## Answer {-}

```{r }
packetTimes = unlist(packetTimes, use.names = FALSE);

overt_interpacket_delays = packetTimes[2:length(packetTimes)] - packetTimes[1:(length(packetTimes)-1)]
covert_interpacket_delays = modifiedPacketStream[2:length(modifiedPacketStream)] - modifiedPacketStream[1:(length(modifiedPacketStream)-1)]
hist(overt_interpacket_delays, main="Overt Packet Stream Inter-Packet Delay Histogram");
hist(covert_interpacket_delays, main="Covert Packet Stream Inter-Packet Delay Histogram")
```

Eve will definitely be suspicious, because the overt interpacket delays seems to follow an exponential distribution as opposed to the covert interpacket delays which takes on one of two values only. If Eve were to see this, Eve would definitely be suspicious.

## 

Alice and Bob decide to use the following modulation scheme. Let $m$, $min$, and $max$ denote the median, min, and max of the inter-packet delay of the overt packet stream. If Alice needs to send a $0$ she randomly generates a delay between $min$ and $m$. If she want to send a $1$ she randomly generates a delay between $m$ and $max$. First, compute $m$, $min$,  and $max$. Next, modify the code  in  Question 1, to generate  the packet stream that contains  the secret message.


## Answer {-}

```{r }
#  calculate median, min, and max of overt inner packet delays
m = mean(packetTimes[2:length(packetTimes)] - packetTimes[1:(length(packetTimes)-1)])
minimum = min(packetTimes[2:length(packetTimes)] - packetTimes[1:(length(packetTimes)-1)])
maximum = max(packetTimes[2:length(packetTimes)] - packetTimes[1:(length(packetTimes)-1)])

print("Median: ")
print(m)
print("Minimum: ")
print(minimum)
print("Maximum: ")
print(maximum)

modifiedPacketStream2 = numeric(8*nchar(secretMsg)+1);
time = 0;
modifiedPacketStream2[1] = time;

for (i in 1:(length(binaryNumbers))) {
  time = time + if(binaryNumbers[i] == 0) runif(1, minimum, m) else runif(1, m, maximum);
  modifiedPacketStream2[i+1] = time;
}

print(modifiedPacketStream2);

```



## 

Plot the histogram  of the inter-packet  delays of the overt packet stream and that of the new covert packet stream. Do you think Eve will be suspicious?
  
## Answer {-}

```{r }
packetTimes = unlist(packetTimes, use.names = FALSE);

overt_interpacket_delays = packetTimes[2:length(packetTimes)] - packetTimes[1:(length(packetTimes)-1)]
covert_interpacket_delays = modifiedPacketStream2[2:length(modifiedPacketStream2)] - modifiedPacketStream2[1:(length(modifiedPacketStream2)-1)]
hist(overt_interpacket_delays, main="Overt Packet Stream Inter-Packet Delay Histogram");
hist(covert_interpacket_delays, main="Covert Packet Stream Inter-Packet Delay Histogram")

```

Eve will definitely be less suspicious because the frequency of the new covert inter packet delays are somewhat evenly distributed between various inter packet delay values while the previous histogram showed inter packet delays of only 0.25 and 0.75. Similar to the histogram of the overt inter packet delays, most of the covert inter packet delays are between 0 - 0.1. Since the histogram of the covert inter packet delays is much similar to the histogram of the overt inter packet delays, Eve will not be suspicious.

##  

Answer the following questions briefly (in 1 or 2 sentences)

1. How can you improve upon the method in Question 3? 

   \textbf{Answer: } Instead of generating equally random delays between the min, median, and max, Alice could have generated random delays based on an exponential distribution that followed the overt inter packet distribution. From there, she could adjust/scale the min, median, and max values to determine how to send a 1 and 0 so the histogram of the covert inter packet delays almost replicates the histogram of the overt inter packet delays.

2. We assumed the Alice will buffer up the packets and we mentioned that it was unrealistic. Why?
 
   \textbf{Answer: } As mentioned before, it's unrealistic for Alice to buffer up her own packets in a Skype call because the quality of the Skype call in terms of both audio and video would be severely compromised. A constant stream of packets must be sent without them being buffered to maintain good performance during the Skype call. Even then, Skype may be programmed to disconnect the call if no packets are being sent over a certain amount of time. Lastly, Alice may not have total control of the time between her inter packet delays.

3. We have assumed that the network does not alter the inter-packet delays. What would be the problem if it did? Can you suggest methods to mitigate the effect of the changes of the inter-packet delay (noise)?

   \textbf{Answer: } If the network altered the inter-packet delays, then the hidden message could be obstructed by sending false 1's or 0's. This could be mitigated by checking how the network affects the covert inter packet delays then adjusting the values used to determine sending 1's and 0's so the noise is accounted for when calculating covert inter packet delays.

# Implementation


## 
For  buffer size $B=20$ we want to find out the probability of overflow  and  underflow, when the IPD follows the Exponential with $\lambda =1$ pkts/sec and $i=2, 6, 10, 14, 18$. Use message size $m = 16, 32$ bits. Tabulate the results. Remember that to determine the probability you need to run multiple (say 500) experiments for each parameter, i.e., for $B = 20, m = 16, i = 2$ run $500$ experiments  and determine the probability of overflow and underflow. Similarly for other values of $i$ and $m$.

## Answer {-}


```{r}
# return values: -1 = underflow, 0 = success, 1 = overflow
expSimulation <- function(maxBufferSize, currentBufferSize, messageSize) {
  median = log(2); # by default log -> ln
  
  # create message.
  bits = sample(0:1, messageSize, replace=T);
  
  # create source packet times.
  overt_packet_times = numeric(200);
  overt_packet_times[1] = 0;
  for (i in 2:200) {
    overt_packet_times[i] = overt_packet_times[i-1] + rexp(1);
  }
  
  time = 0;
  
  time = 0;
  for (bit in bits) {
    # figure out the delay needed for the current packet.
    delay = -1;
    while(T) {
      delay = rexp(1);
      
      if (bit == 0 && delay < median) {
        break;
      }
      
      if (bit == 1 && delay >= median) {
        break
      }
    }
    
    # send the packet
    currentBufferSize = currentBufferSize - 1;
    
    # buffer in arriving packets.
    currentBufferSize = currentBufferSize + length(overt_packet_times[(overt_packet_times > time) & (overt_packet_times <= (time+delay))]);
    
    # update time
    time = time + delay;
    
    # check for overflow.
    if (currentBufferSize > maxBufferSize) {
      return(1);
    }
    
    # check for underflow (underflow only occurs if buffer is empty)
    if (currentBufferSize == 0 && bit == 0) {
      # if no packets arrive, the message will get corrupted.
      timeRemainingTillUnderflow = if(bit == 0) median-delay;
      
      arrivingPackets = length(overt_packet_times[(overt_packet_times > time) & (overt_packet_times <= (time+timeRemainingTillUnderflow))]);
      
      # no packets arrived during this time.
      if (arrivingPackets == 0) {
        return(-1);
      }
    }
  }
  return(0);
}

```

```{r}
initBufferSizes = c(2,6,10,14,18);
m = c(16,32);

underflows16 = numeric(5);
underflows32 = numeric(5);
overflows16 = numeric(5);
overflows32 = numeric(5);

for (i in 1:length(initBufferSizes)) {
  overflows = 0;
  underflows = 0;
  for (j in 1:500) {
    value = expSimulation(20,initBufferSizes[i],16);
    if (value == 1) {
      overflows = overflows + 1;
    }
    else if (value == -1) {
      underflows = underflows + 1;
    }
  }
  underflows16[i] = underflows / 500;
  overflows16[i] = overflows / 500;
}

for (i in 1:length(initBufferSizes)) {
  overflows = 0;
  underflows = 0;
  for (j in 1:500) {
    value = expSimulation(20,initBufferSizes[i],32);
    if (value == 1) {
      overflows = overflows + 1;
    }
    else if (value == -1) {
      underflows = underflows + 1;
    }
  }
  underflows32[i] = underflows / 500;
  overflows32[i] = overflows / 500;
}

ans = matrix(c(underflows16,overflows16, underflows32,overflows32), nrow=4, ncol=5, byrow = T);
dimnames(ans) = list(c("underflow % (m=16)", "overflow  % (m=16)", "underflow % (m=32)", "overflow  % (m=32)"),c("i=2", "i=6", "i=10", "i=14", "i=18"))
print("Exponential Distribution")
print(ans)
```

## 
For  buffer size $B=20$ we want to find out the probability of overflow  and  underflow, when the IPD follows the Uniform distribution in the range (0,1)  and $i=2, 6, 10, 14, 18$. Use message size $m = 16, 32$ bits.  Tabulate the results. Remember that to determine the probability you need to run multiple (say 500) experiments for each parameter, i.e., for $B = 20, m = 16, i = 2$ run $500$ experiments  and determine the probability of overflow and underflow. Similarly for other values of $i$ and $m$.

## Answer {-}


```{r}
# return values: -1 = underflow, 0 = success, 1 = overflow
unifSimulation <- function(maxBufferSize, currentBufferSize, messageSize) {
  median = 0.5;
  
  # create message.
  bits = sample(0:1, messageSize, replace=T);
  
  # create source packet times.
  overt_packet_times = numeric(200);
  overt_packet_times[1] = 0;
  for (i in 2:200) {
    overt_packet_times[i] = overt_packet_times[i-1] + runif(1);
  }
  
  time = 0;
  
  time = 0;
  for (bit in bits) {
    # figure out the delay needed for the current packet.
    delay = if(bit == 1) runif(1, 0.5,1) else runif(1,0,0.5);
    
    # send the packet
    currentBufferSize = currentBufferSize - 1;
    
    # buffer in arriving packets.
    currentBufferSize = currentBufferSize + length(overt_packet_times[(overt_packet_times > time) & (overt_packet_times <= (time+delay))]);
    
    # update time
    time = time + delay;
    
    # check for overflow.
    if (currentBufferSize > maxBufferSize) {
      return(1);
    }
    
    # check for underflow (underflow only occurs if buffer is empty)
    if (currentBufferSize == 0) {
      timeRemainingTillUnderflow = if(bit == 1) 1-delay else median-delay;
      
      arrivingPackets = length(overt_packet_times[(overt_packet_times > time) & (overt_packet_times <= (time+timeRemainingTillUnderflow))]);
      
      # no packets arrived during this time.
      if (arrivingPackets == 0) {
        return(-1);
      }
    }
  }
  
  return(0);
}
```

```{r}
# tabulate results and compute the probabilities
initBufferSizes = c(2,6,10,14,18);
m = c(16,32);

underflows16 = numeric(5);
underflows32 = numeric(5);
overflows16 = numeric(5);
overflows32 = numeric(5);

for (i in 1:length(initBufferSizes)) {
  overflows = 0;
  underflows = 0;
  for (j in 1:500) {
    value = unifSimulation(20,initBufferSizes[i],16);
    if (value == 1) {
      overflows = overflows + 1;
    }
    else if (value == -1) {
      underflows = underflows + 1;
    }
  }
  underflows16[i] = underflows / 500;
  overflows16[i] = overflows / 500;
}

for (i in 1:length(initBufferSizes)) {
  overflows = 0;
  underflows = 0;
  for (j in 1:500) {
    value = unifSimulation(20,initBufferSizes[i],32);
    if (value == 1) {
      overflows = overflows + 1;
    }
    else if (value == -1) {
      underflows = underflows + 1;
    }
  }
  underflows32[i] = underflows / 500;
  overflows32[i] = overflows / 500;
}

ans = matrix(c(underflows16,overflows16, underflows32,overflows32), nrow=4, ncol=5, byrow = T);
dimnames(ans) = list(c("underflow % (m=16)", "overflow  % (m=16)", "underflow % (m=32)", "overflow  % (m=32)"),c("i=2", "i=6", "i=10", "i=14", "i=18"))
print("Uniform Distribution")
print(ans)
```

##  
Propose methods to deal with buffer overflow and underflow.

## Answer {-}
Choose the Uniform or Exponential distribution based on how high the success rate is.

Scale up the time remaining until underflow happens so a greater time window can catch more packets.

Scale down the generated delays between packets so less packets are added to the buffer between a given time and a delay.

Have a bigger buffer size.

